---
title: "TSA Forecast Competition Final Report"
subtitle: "https://github.com/maevegr/GualtieriReed_Kasten_Zhang_ENV797_TSA_ForecastCompetition_S26"
author: "Maeve Gualtieri-Reed and Sadie Kasten"
date: "2026-04-24"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=F}
library(here)
library(readxl)
library(dplyr)
library(tidyverse)
library(forecast)
library(cowplot)
library(smooth)
library(writexl)
library(tinytex)
library(kableExtra)
library(knitr)
```

# Introduction 

This report forecasts July 2011 electricity load using daily electricity load, temperature, and humidity data from January 1, 2005 to June 30, 2011. It uses a training window of January 1, 2005 to June 30, 2010 to forecast and test models compared to the test window of June 30, 2010. 

Full code and testing done by each report contributor can be found in the GitHub repository: https://github.com/maevegr/GualtieriReed_Kasten_Zhang_ENV797_TSA_ForecastCompetition_S26

ChatGPT was used to learn how to add exogenous variables to ARIMA + Fourier and Neural Network. It was also used for general debugging.

# Methods

The load data was uploaded and converted to daily data. It was converted to a time series using msts().

```{r echo=TRUE, message=FALSE, warning=FALSE}
load <- read_excel("load.xlsx") %>% 
  mutate(daily = rowMeans(across(3:26), na.rm = TRUE)) %>% 
  mutate(date = ymd(date)) %>% 
  select(date,daily)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
msts_load <- msts(load$daily, 
                seasonal.periods =c(7,365.25), #7 = weekly, 365.25 = yearly 
                start=c(2005,1,1))
```

A training and testing section of the data were separated out in order to estimate model suitability before uploading into Kaggle

```{r echo=TRUE, message=FALSE, warning=FALSE}

ts_load_train <- subset(msts_load,
                       end = length(msts_load)-366)

ts_load_test <- subset(msts_load, start = length(msts_load)- 365,
                       end = length(msts_load)-335)

```

The data was examined to determine what types of models might be most suitable for the data. It is clear there are strong multiple seasonalities which indicates ARIMA + Fourier and TBATS will be good candidates. ETS + STL may not be a great model because it seems that the seasonality is not perfectly stable as indicated by the persistence of some kind of pattern in the decomposition residuals. Neural network will also be explored to see if there are non-linear patterns that are better predicted through a learning model.

```{r echo=TRUE, message=FALSE, warning=FALSE, fig.cap="Plot of Raw Data"}
autoplot(msts_load)
```

```{r echo=TRUE, message=FALSE, warning=FALSE, fig.cap = "Decomposition"}
decomp <- msts_load %>% mstl() 
autoplot(decomp)
```

Exogenous variables were loaded and turned into vectors. To be used successfully in the model forecast, forecasts of the exogenous variables also had to be constructed. This was done with the TBATS model after comparison between ARIMA and TBATS on one forecasting model proved better results via the TBATS forecast of the variables.

```{r echo=TRUE, message=FALSE, warning=FALSE}
temp_df <- read_excel("temperature.xlsx") %>% 
  mutate(hourly_temp = rowMeans(across(3:26), na.rm = TRUE)) %>% 
  mutate(date = ymd(date)) %>% 
  group_by(date) %>%
  summarise(daily = mean(as.numeric(hourly_temp), na.rm = TRUE)) %>%
  select(daily) %>% 
  na.omit(temp)

temp <- temp_df$daily #vector

temp_train <- temp[1:(length(temp) - 366)]
temp_test <- temp[(length(temp) - 365):(length(temp) - 335)]
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
humidity_df <- read_excel("relative_humidity.xlsx") %>% 
  mutate(hourly_humidity = rowMeans(across(3:26), na.rm = TRUE)) %>% 
  mutate(date = ymd(date)) %>% 
  group_by(date) %>%
  summarise(daily = mean(as.numeric(hourly_humidity), na.rm = TRUE)) %>%
  select(daily) %>%
  na.omit(humidity)

humidity <- humidity_df$daily #vector

humidity_train <- humidity[1:(length(humidity) - 366)]
humidity_test <- humidity[(length(humidity) - 365):(length(humidity) - 335)]
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
temp_July <- forecast(tbats(temp), h = 31) #forecasting future 31 from the vector
humidity_July <- forecast(tbats(humidity), h = 31) #forecasting future 31 from the vector 
```

Exogenous variables were tested via correlation tests to determine if both should be included in the forecasting models. Based on correlation tests and testing models with and without each variable, both exogenous variables were used, if used in a forecasting model.

```{r echo=TRUE, message=FALSE, warning=FALSE}
cor(msts_load, temp)
cor(msts_load, humidity)
cor(temp,humidity)
```

Model selection was determined by which were forecasting models would be best suited to handle load data and exogenous variables if used. The TBATS, ARIMA + fourier terms, ETS, and Neural Network forecasting models were all selected as testing candidates.

# Results

**The top 5 models are:**

1)  ARIMA + Fourier (K = 2,12) with temperature and humidity as exogenous variables
2)  ARIMA + Fourier (K = 2,10) with temperature and humidity as exogenous variables
3)  TBATS
4)  Neural Network (K = 2,1) with temperature and humidity as exogenous variables
5)  Neural Network (K = 2,12) with temperature and humidity as exogenous variables

Additional models, including ETS + STL, ARIMA + Fourier and NN without exogenous variables, as well as multiple attempts with different K values were tested. ETS+STL resulted in a MAPE score of 0.1330, which is better than the seasonal naive, but not as good as the ARIMA, TBATS, or NN.

## TBATS 
The first of the five top models that performed best against the actual data for July 2011 was created using the TBATS forecasting model. This model was trained using the simply the load data and forecast for the month of July. The MAPE score of this forecasting model was 0.0998. The forecast result is plotted below.

```{r echo=TRUE, message=FALSE, warning=FALSE}
#TBATS Training Forecast & Test Window

TBATS_train <- tbats(ts_load_train)

TBATS_train_forecast <- forecast(TBATS_train, h = 31)

accuracy(TBATS_train_forecast$mean, ts_load_test)
```

```{r echo=TRUE, message=FALSE, warning=FALSE, fig.cap="TBATS Forecast"}
#TBATS July 2011 forecast
TBATS_model <- tbats(msts_load)

TBATS_forecast <- forecast(TBATS_model, h=31)

plot(TBATS_forecast)

TBATS_data<- data.frame(
  date = seq(as.Date("2011-07-01"), length.out = 31, by = "day"),
  value = TBATS_forecast$mean
)

#write_csv(TBATS_data, 'Submissions/TBATS_data.csv')
```

## Neural Network 
Below are the two Neural Network models that also ranked among our top 5 forecasts based on actual data.

The first Neural Network model includes exogenous variables for temperature and humidity as well as exogenous K values 2 and 12 after testing multiple versions of K value options. The MAPE score for this forecast was 0.1169 after optimization.

```{r echo=TRUE, message=FALSE, warning=FALSE}

# NN K = 2,12 Training Forecast & Test Window

#start with train window
temp<-temp_train
humidity<-humidity_train

#training model
NN_train1 <- nnetar(ts_load_train,
                 p=10,
                 P=0,
                 xreg= cbind(fourier(ts_load_train, K=c(2,12)), temp, humidity))

#change to test window
temp <- temp_test
humidity <- humidity_test

#forecast with train model
NN_train_forecast1 <- forecast(NN_train1,
                   xreg= cbind(fourier(ts_load_train, K = c(2,12), h = 31),
                               temp,   
                               humidity))

accuracy(NN_train_forecast1$mean, ts_load_test)

```

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.cap="Neural Network, Version 1"}
## NN K = 2,12 July 2011 forecast
NN_4 <- nnetar(msts_load,
                 p=10,
                 P=0,
                 xreg= cbind(fourier(msts_load, K=c(2,12)), temp, humidity))

NN_4_forecast <- forecast(NN_4,
                   xreg= cbind(fourier(msts_load, K = c(2,12), h = 31),
                               as.numeric(temp_July$mean),   
                               as.numeric(humidity_July$mean)))  

plot(NN_4_forecast)

NN_4_data<- data.frame(
  date = seq(as.Date("2011-07-01"), length.out = 31, by = "day"),
  value = NN_4_forecast$mean
)

#write_csv(NN_4_data, 'Submissions/NN_4_data.csv')
```

The second Neural Network model also includes exogenous variables for temperature and humidity as well as exogenous K values 2 and 1 after testing multiple versions of K value options. The MAPE score for this forecast was 0.1134 after optimization.

```{r echo=TRUE, message=FALSE, warning=FALSE}
## NN K = 2,12 Training Forecast & Test Window

#start with train window
temp<-temp_train
humidity<-humidity_train

#training model
NN_train2 <- nnetar(ts_load_train,
                 p=10,
                 P=1,
                 xreg= cbind(fourier(ts_load_train, K=c(2,1)), temp, humidity))

#change to test window
temp <- temp_test
humidity <- humidity_test

#forecast with train model
NN_train_forecast2 <- forecast(NN_train2,
                   xreg= cbind(fourier(ts_load_train, K = c(2,1), h = 31),
                               temp,   
                               humidity))

accuracy(NN_train_forecast2$mean, ts_load_test)
```

```{r echo=TRUE, message=FALSE, warning=FALSE, fig.cap="Neural Network, Version 2"}

NN_6 <- nnetar(msts_load,
                 p=10,
                 P=1,
                 xreg= cbind(fourier(msts_load, K=c(2,1)), temp, humidity))

NN_6_forecast <- forecast(NN_6,
                   xreg= cbind(fourier(msts_load, K = c(2,1), h = 31),
                               as.numeric(temp_July$mean),   
                               as.numeric(humidity_July$mean)))
plot(NN_6_forecast)

NN_6_data<- data.frame(
  date = seq(as.Date("2011-07-01"), length.out = 31, by = "day"),
  value = NN_6_forecast$mean
)

#write_csv(NN_6_data, 'Submissions/NN_6_data.csv')
```

## ARIMA + Fourier Terms 
Similar to the Neural Network models, the ARIMA models that ranked best compared to the actual data also included exogenous variables for temperature and humidity.

Our highest performing ARIMA model, and model overall, included fourier terms where K= 2 and 12 and temp and humidity forecasts were included. The MAPE score for this forecast was 0.0965 after optimization.

Initially, K = 2,12 was used as a baseline estimate for the best K values. Looking at the test window of July 2010, this model does not perform very well compared to the other models. However, using the full dataset to train and forecast July 2011, this model is our best performing with the lowest MAPE.
```{r echo=TRUE, message=FALSE, warning=FALSE}
## ARIMA +FOURIER (K = 2,12) Training Forecast & Test Window

#start with train window
temp<-temp_train
humidity<-humidity_train

ARIMAF_train1 <- auto.arima(ts_load_train, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg= cbind(fourier(ts_load_train, K=c(2,12)), temp, humidity))

#change to test window
temp <- temp_test
humidity <- humidity_test

ARIMAF_train_forecast1<- forecast(ARIMAF_train1,
                        xreg= cbind(fourier(ts_load_train, K = c(2,12), h = 31),
                               temp,   
                               humidity), h=31)

accuracy(ARIMAF_train_forecast1$mean, ts_load_test)

```

```{r eecho=TRUE, message=FALSE, warning=FALSE, fig.cap="ARIMA+Fourier, Version 1"}

ARIMAF_1 <- auto.arima(msts_load, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg= cbind(fourier(msts_load, K=c(2,12)), temp, humidity))

ARIMAF_1_forecast<- forecast(ARIMAF_1,
                        xreg= cbind(fourier(msts_load, K = c(2,12), h = 31),
                               as.numeric(temp_July$mean),   
                               as.numeric(humidity_July$mean)), h=31)

plot(ARIMAF_1_forecast)

ARIMAF_1_data<- data.frame(
  date = seq(as.Date("2011-07-01"), length.out = 31, by = "day"),
  value = ARIMAF_1_forecast$mean
)

#write_csv(ARIMAF_1_data, 'Submissions/ARIMAF_1_data.csv')
```

A loop was run for k values between 1 and 3 for the weekly seasonality and between 8 and 15 for the annual seasonality. K = 2,10 yielded the best MAPE score and all AIC values were fairly similar. It did not however beat out the K = 2,12 after uploading to Kaggle. Therefore, the test window wasn't a perfect estimate of the actual forecasting window for model fitting. 

```{r echo=TRUE, message=FALSE, warning=FALSE}

## ARIMA +FOURIER (K = 2,10) Training Forecast & Test Window

#start with train window
temp<-temp_train
humidity<-humidity_train

ARIMAF_train2 <- auto.arima(ts_load_train, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg= cbind(fourier(ts_load_train, K=c(2,10)), temp, humidity))

#change to test window
temp <- temp_test
humidity <- humidity_test

ARIMAF_train_forecast2<- forecast(ARIMAF_train2,
                        xreg= cbind(fourier(ts_load_train, K = c(2,10), h = 31),
                               temp,   
                               humidity), h=31)

accuracy(ARIMAF_train_forecast2$mean, ts_load_test)
```

```{r echo=TRUE, message=FALSE, warning=FALSE, fig.cap="ARIMA+Fourier, Version 2"}

ARIMAF_2 <- auto.arima(msts_load, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg= cbind(fourier(msts_load, K=c(2,10)), temp, humidity))

ARIMAF_2_forecast<- forecast(ARIMAF_2,
                        xreg= cbind(fourier(msts_load, K = c(2,10), h = 31),
                               as.numeric(temp_July$mean),   
                               as.numeric(humidity_July$mean)), h=31)

plot(ARIMAF_2_forecast)

ARIMAF_2_data<- data.frame(
  date = seq(as.Date("2011-07-01"), length.out = 31, by = "day"),
  value = ARIMAF_2_forecast$mean
)

#write_csv(ARIMAF_2_data, 'Submissions/ARIMAF_1_data.csv')

```

\newpage 

## Summary of all models

```{r message=FALSE, echo=FALSE, warning=FALSE}

TBATS_score <- round(accuracy(TBATS_train_forecast$mean, ts_load_test),2)
NN1_score <- round(accuracy(NN_train_forecast1$mean, ts_load_test),2)
NN2_score <- round(accuracy(NN_train_forecast2$mean, ts_load_test),2)
ARIMAF1_score <- round(accuracy(ARIMAF_train_forecast1$mean, ts_load_test),2)
ARIMAF2_score <- round(accuracy(ARIMAF_train_forecast2$mean, ts_load_test),2)

Names <- c("TBATS","NN K=2,12", "NN K=2,1", "ARIMAF K=2,12", "ARIMAF K=2,10")
July2011_MAPE <- c("9.98", "11.34", "11.38", "9.65", "9.93")

summary <- data.frame() %>% 
  rbind(TBATS_score, NN1_score, NN2_score, ARIMAF1_score, ARIMAF2_score) %>% 
  mutate(Name = Names) %>% 
  mutate(July2011MAPE = July2011_MAPE) %>% 
  select(Name, July2011MAPE, ME,RMSE,MAE,MPE,MAPE,ACF1,`Theil's U`)

kbl(summary, 
      caption = "Forecast Accuracy for Electricy Load",
      digits = array(5,ncol(summary))) %>%
  kable_styling(full_width = FALSE, position = "center", latex_options = "hold_position")
  
```

```{r message=FALSE, echo=FALSE, warning=FALSE, fig.cap="All Forecasts"}
all_forecasts<- autoplot(ts_load_test)+
  autolayer(TBATS_train_forecast, series = "TBATS", PI = FALSE)+
  autolayer(NN_train_forecast1, series = "NN (K = 2,12)", PI = FALSE)+
  autolayer(NN_train_forecast2, series = "NN (K = 2,1)", PI = FALSE)+
  autolayer(ARIMAF_train_forecast1, series = "ARIMA+Fourier (K = 2,12)", PI = FALSE)+
  autolayer(ARIMAF_train_forecast2, series = "ARIMA+Fourier (K = 2,10)", PI = FALSE)+
  ylab("Load")+
  xlab("July 2010")

print(all_forecasts)
```

# Analysis and Conclusions

Using the training set up until June 30th, 2010 and comparing each models forecast for July 2010 to the actual data, the neural network model with terms K = 2,12 was the best fit with the lowest MAPE score. However, after using these models to forecast July 2011 using all data and uploading the forecast to Kaggle, the ARIMA + Fourier with terms K = 2,12 yielded the best MAPE score. This indicates that while training and testing models can help determine which model types work best, testing windows may not always estimate the best forecasting model to use. While the Neural-Network models do the best, none of the models adequately predicted the extremes seen in July 2010. The data should be more deeply explored to determine if July 2010 was an anomaly year or if it displayed a level shift or other major change that would have significantly helped predict July 2011. 

# Unused Models
Code for models that did not make the top 5 models are not included in this report but can be found in the corresponding .rmd file in GitHub. Model testing code by team member can be found in the Testing folder in GitHub.
```{r eval=FALSE, include=FALSE}
#Unused Model Code
#ARIMA 212
ARIMA_212 <- auto.arima(ts_load_train, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_load_train, 
                                          K=c(2,12))
                             )

#Forecast with ARIMA fit
#also need to specify h for fourier terms
ARIMA_212_forecast <- forecast(ARIMA_212,
                           xreg=fourier(ts_load_train,
                                        K=c(2,12),
                                        h=31),
                           h=31
                           ) 

ARIMA_212_scores <- accuracy(ARIMA_212_forecast$mean, ts_load_test)

ARIMA_212_data<- data.frame(
  date = seq(as.Date("2011-07-01"), length.out = 31, by = "day"),
  value = ARIMA_212_forecast$mean
)

write_csv(ARIMA_212_data, 'Submissions/ARIMAfour1.csv')
```

```{r eval=FALSE, include=FALSE}

#ARIMA 3,8 with humidity and temp also modeled with ARIMA 2,10
ARIMA_temp <- auto.arima(ts_temp_full, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_temp_full, 
                                          K=c(2,10))
                             )

ARIMA_humid <- auto.arima(ts_humid_full, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_humid_full, 
                                          K=c(2,10))
                             )

ARIMA_temp_forecast <- forecast(ARIMA_temp,
                           xreg=fourier(ts_temp_full, 
                                           K=c(2,12),
                                        h=31),
                           h=31
                             )

ARIMA_humid_forecast <- forecast(ARIMA_humid,
                           xreg=fourier(ts_humid_full, 
                                           K=c(2,12),
                                        h=31),
                           h=31
                             )

temp_forecast <- data.frame(avg_temp = ARIMA_temp_forecast$mean)
humid_forecast <- data.frame(avg_humidity = ARIMA_humid_forecast$mean)

fourier_terms_forecast <- fourier(ts_load_full, K=c(3,8), h= 31)
xreg_data_forecast <- cbind(temp_forecast,humid_forecast,fourier_terms_forecast) %>% 
  as.matrix()


ARIMA_exog38_forecast <- forecast(ARIMA_exog38,
                           xreg=xreg_data_forecast,
                           h=31
                           ) 

ARIMA_exog38_data<- data.frame(
  date = seq(as.Date("2011-07-01"), length.out = 31, by = "day"),
  value = ARIMA_exog38_forecast$mean
)

write_csv(ARIMA_exog38_data, 'ARIMA_exog38_data.csv')
```

```{r eval=FALSE, include=FALSE}
#ARIMA k = 2,10 with temp and humidity TBATS
fourier_terms <- fourier(ts_load_full, K=c(2,10))
xreg_data <- left_join(temp_full, humidity_full, by = "date") %>% 
  head(-1) %>% 
  select(avg_temp, avg_humidity) %>% 
  cbind(fourier_terms) %>% 
  as.matrix()


ARIMA_exog210 <- auto.arima(ts_load_full, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=xreg_data
                             )
temp_clean <- na.interp(temp_full$avg_temp)
temp_tbats <- tbats(temp_clean)
temp_tbats_forecast <- forecast(temp_tbats, h = 31)

humid_clean <- na.interp(humidity_full$avg_humidity)
humid_tbats <- tbats(humid_clean)
humid_tbats_forecast <- forecast(humid_tbats, h = 31)

temp_forecast <- data.frame(avg_temp = temp_tbats_forecast$mean)
humid_forecast <- data.frame(avg_humidity = humid_tbats_forecast$mean)

fourier_terms_forecast <- fourier(ts_load_full, K=c(2,10), h= 31)
xreg_data_forecast <- cbind(temp_forecast,humid_forecast,fourier_terms_forecast) %>% 
  as.matrix()


ARIMA_exog210_forecast <- forecast(ARIMA_exog210,
                           xreg=xreg_data_forecast,
                           h=31
                           ) 

ARIMA_exog210tbats_data<- data.frame(
  date = seq(as.Date("2011-07-01"), length.out = 31, by = "day"),
  value = ARIMA_exog210_forecast$mean
)

write_csv(ARIMA_exog210tbats_data, 'ARIMA_exog210tbats_data.csv')
```

```{r eval=FALSE, include=FALSE}
temp_clean <- na.interp(temp_train$avg_temp)
temp_clean_df <- data.frame(avg_temp = temp_clean)
temp_tbats <- tbats(temp_clean)
temp_tbats_forecast <- forecast(temp_tbats, h = 31)


humid_clean <- na.interp(humidity_train$avg_humidity)
humid_clean_df <- data.frame(avg_humidity = humid_clean)
humid_tbats <- tbats(humid_clean)
humid_tbats_forecast <- forecast(humid_tbats, h = 31)


temp_forecast <- data.frame(avg_temp = temp_tbats_forecast$mean)
humid_forecast <- data.frame(avg_humidity = humid_tbats_forecast$mean)

results <- data.frame()
for (k1 in 1:3) {
  for (k2 in 8:15) {
    xreg_fit <- cbind(temp_clean_df,humid_clean_df,
                                   fourier(ts_load_train, K = c(k1,k2))) %>% 
      as.matrix()
    fit <- auto.arima(ts_load_train,
                      xreg = xreg_fit,
                      seasonal = FALSE)
    xreg_forecast <- cbind(temp_forecast, humid_forecast, fourier(ts_load_train,
                                        K=c(k1,k2),
                                        h=31)) %>% 
      as.matrix()
    forecast <- forecast(fit,
                           xreg=xreg_forecast,
                           h=31
                           ) 
    AIC<- AIC(fit)
    scores<- accuracy(forecast$mean, ts_load_test)
    results <- rbind(results,
                     data.frame(K1=k1, K2=k2, scores, AIC = as.numeric(AIC)))
  }
}

### Run a loop for no exog

results <- data.frame()
for (k1 in 1:3) {
  for (k2 in 8:15) {
    fit <- auto.arima(ts_load_train,
                      xreg = fourier(ts_load_train, K = c(k1, k2)),
                      seasonal = FALSE)
    forecast <- forecast(fit,
                           xreg=fourier(ts_load_train,
                                        K=c(k1,k2),
                                        h=31),
                           h=31
                           ) 
    scores<- accuracy(forecast$mean, ts_load_test)
    results <- rbind(results,
                     data.frame(K1=k1, K2=k2, scores))
  }
}
```

```{r eval=FALSE, include=FALSE}
# Top ARIMA models (no exog) AIC values
ARIMA_110 <- auto.arima(ts_load_train, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_load_train, 
                                          K=c(1,10))
                             )

ARIMA_18 <- auto.arima(ts_load_train, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_load_train, 
                                          K=c(1,8))
                             )

ARIMA_111 <- auto.arima(ts_load_train, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_load_train, 
                                          K=c(1,11))
                             )

ARIMA_310 <- auto.arima(ts_load_train, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_load_train, 
                                          K=c(3,10))
                             )

ARIMA_38 <- auto.arima(ts_load_train, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_load_train, 
                                          K=c(3,8))
                             )

ARIMA_311 <- auto.arima(ts_load_train, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_load_train, 
                                          K=c(3,11))
                             )


AIC <- data.frame() %>% 
  rbind(data.frame(Model=110, AIC = AIC(ARIMA_110))) %>% 
  rbind(data.frame(Model=18, AIC = AIC(ARIMA_18))) %>% 
  rbind(data.frame(Model=111, AIC = AIC(ARIMA_111))) %>% 
  rbind(data.frame(Model=310, AIC = AIC(ARIMA_310))) %>% 
  rbind(data.frame(Model=38, AIC = AIC(ARIMA_38))) %>% 
  rbind(data.frame(Model=311, AIC = AIC(ARIMA_311)))
```

```{r eval=FALSE, include=FALSE}
#Fit and forecast STL + ETS model to data
ETS_train1 <-  stlf(ts_load_train,h=31)

#Plot foresting results
autoplot(ETS_train1) + ylab("Load")

#Plot model + observed data
autoplot(ts_load_full) +
  autolayer(ETS_train1, series="STL + ETS",PI=FALSE) +
  ylab("Load")

ETS_train_scores<- accuracy(ETS_train1$mean,ts_load_full)  

ETS_fit1<- stlf(ts_load_full,h=31)
ETS_output1<- data.frame(
  date = seq(as.Date("2011-07-01"), length.out = 31, by = "day"),
  value = ETS_fit1$mean
)

write_csv(ETS_output1, 'Submissions/ETS1.csv')
```

```{r eval=FALSE, include=FALSE}
#basic TS
ts_load <- ts(load$daily, start=c(2005,1,1))
#autoarima
arima_attempt1 <- auto.arima(ts_load, ic=c("aic"))
#forecast based on the auto arima model
arima_forecast1 <- forecast(arima_attempt1, h=31)

#create dataframe /prep for export 
arima1_df <- data.frame(load = arima_forecast1$mean)
arima1_df <- arima1_df %>% 
  add_column(date = seq(as.Date("2011-07-01"), by = "day", length.out = nrow(arima1_df)),.before = 1)

#export as an excel sheet
write_csv(arima1_df, 'Submissions/arima1.csv')
```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#neural network, p=1 P=0, K=2,12
#test diff p and P values, also can add xreg with Fourier term for multiple seasonality

NN_1 <- nnetar(msts_load,
                 p=1,
                 P=0,
                 xreg=fourier(msts_load, K=c(2,12)))

NN_1_forecast <- forecast(NN_1, 
                   h=31,
                   xreg=fourier(msts_load, K=c(2,12),h=31))

#create dataframe /prep for export 
NN_1_df <- data.frame(load = NN_1_forecast$mean)
NN_1_df <- NN_1_df %>% 
  add_column(date = seq(as.Date("2011-07-01"), by = "day", length.out = nrow(NN_1_df)),.before = 1)

#export as an excel sheet
write_csv(NN_1_df, 'Submissions/NN_1.csv')
```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#test diff p and P values, also can add xreg with Fourier term for multiple seasonality
#neural network, p=7 P=0, K=2,12
#internet suggested p=7 / a higher p would work better. maybe.

NN_2 <- nnetar(msts_load,
                 p=7,
                 P=0,
                 xreg=fourier(msts_load, K=c(2,12)))

NN_2_forecast <- forecast(NN_2, 
                   h=31,
                   xreg=fourier(msts_load, K=c(2,12),h=31))

#create dataframe /prep for export 
NN_2_df <- data.frame(load = NN_2_forecast$mean)
NN_2_df <- NN_2_df %>% 
  add_column(date = seq(as.Date("2011-07-01"), by = "day", length.out = nrow(NN_2_df)),.before = 1)

#export as an excel sheet
write_csv(NN_2_df, 'Submissions/NN_2.csv')
```



```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#test diff p and P values, also can add xreg with Fourier term for multiple seasonality
##neural network, p=7 P=0, K=2,6

NN_3 <- nnetar(msts_load,
                 p=7,
                 P=0,
                 xreg=fourier(msts_load, K=c(2,6)))

NN_3_forecast <- forecast(NN_3, 
                   h=31,
                   xreg=fourier(msts_load, K=c(2,6),h=31))

#create dataframe /prep for export 
NN_3_df <- data.frame(load = NN_3_forecast$mean)
NN_3_df <- NN_3_df %>% 
  add_column(date = seq(as.Date("2011-07-01"), by = "day", length.out = nrow(NN_3_df)),.before = 1)

#export as an excel sheet
write_csv(NN_3_df, 'Submissions/NN_3.csv')
```

```{r eval=FALSE, include=FALSE}

#Should add temp and humidity to improve my best model - NN_2!
#neural network, p=10 P=0, K=2,12 + temp and humidity variables
temp_df <- read_excel("temperature.xlsx") %>% 
  mutate(hourly_temp = rowMeans(across(3:26), na.rm = TRUE)) %>% 
  mutate(date = ymd(date)) %>% 
  group_by(date) %>%
  summarise(daily = mean(as.numeric(hourly_temp), na.rm = TRUE)) %>%
  select(daily) %>%
  na.omit(temp)

temp <- temp_df$daily #vector
```

```{r eval=FALSE, include=FALSE}
humidity_df <- read_excel("relative_humidity.xlsx") %>% 
  mutate(hourly_humidity = rowMeans(across(3:26), na.rm = TRUE)) %>% 
  mutate(date = ymd(date)) %>% 
  group_by(date) %>%
  summarise(daily = mean(as.numeric(hourly_humidity), na.rm = TRUE)) %>%
  select(daily) %>%
  na.omit(humidity)

humidity <- humidity_df$daily #vector
```

```{r eval=FALSE, include=FALSE}
autoplot(NN_1_forecast$mean)+
  autolayer(NN_2_forecast$mean)+
  autolayer(NN_3_forecast$mean)+
  autolayer(NN_4_forecast$mean) #forecasting temp and humidity with tbats much better than auto.arima() - similar to the other NN forecasts
#The forecast for NN_4 including temp and humidity was much worse when using auto.arima() to forecast future values. Performed much better when swapping to TBATS - the auto.arima forecast way below the others as it predicted temp to fall off.
```

```{r eval=FALSE, include=FALSE}
cor(msts_load, temp)
cor(msts_load, humidity)
cor(temp,humidity)
```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#trying one more NN model - this time only temp since its stronger correlated to load than humidity.


#neural network, p=7 P=0, K=2,12 + temp variable
NN_5 <- nnetar(msts_load,
                 p=7,
                 P=0,
                 xreg= cbind(fourier(msts_load, K=c(2,12)), temp))

temp_July2 <- forecast(tbats(temp), h = 31) #forecasting future 31 from the vector

NN_5_forecast <- forecast(NN_5,
                   xreg= cbind(fourier(msts_load, K = c(2,12), h = 31),
                               as.numeric(temp_July2$mean)))   

#create dataframe /prep for export 
NN_5_df <- data.frame(load = NN_5_forecast$mean)
NN_5_df <- NN_5_df %>% 
  add_column(date = seq(as.Date("2011-07-01"), by = "day", length.out = nrow(NN_5_df)),.before = 1)

#export as an excel sheet
write_csv(NN_5_df, 'Submissions/NN_5.csv')
```
