---
title: "Sadie"
author: "Sadie Kasten"
date: "2026-03-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Setup
```{r}
library(here)
library(readxl)
library(dplyr)
library(tidyverse)
library(forecast)
library(cowplot)
library(smooth)
library(writexl)
```


# Load in data set

```{r}
load <- read_excel("load.xlsx") %>% 
  mutate(daily = rowMeans(across(3:26), na.rm = TRUE)) %>% 
  mutate(date = ymd(date)) %>% 
  select(date,daily)
```


## Create time series object

```{r}
msts_load <- msts(load$daily, 
                seasonal.periods =c(7,365.25), #7 = weekly, 365.25 = yearly 
                start=c(2005,1,1))
```

#Examine data

```{r}
autoplot(msts_load)
```


#Creating Basic Model

```{r}
#basic TS
ts_load <- ts(load$daily, start=c(2005,1,1))
#autoarima
arima_attempt1 <- auto.arima(ts_load, ic=c("aic"))
#forecast based on the auto arima model
arima_forecast1 <- forecast(arima_attempt1, h=31)

#create dataframe /prep for export 
arima1_df <- data.frame(load = arima_forecast1$mean)
arima1_df <- arima1_df %>% 
  add_column(date = seq(as.Date("2011-07-01"), by = "day", length.out = nrow(arima1_df)),.before = 1)

#export as an excel sheet
write_csv(arima1_df, 'Submissions/arima1.csv')

```


#TBATS model
```{r}
#TBATS for for load no training
TBATS_model <- tbats(msts_load)

TBATS_forecast <- forecast(TBATS_model, h=31)

#create dataframe /prep for export 
TBATS_df <- data.frame(load = TBATS_forecast$mean)
TBATS_df <- TBATS_df %>% 
  add_column(date = seq(as.Date("2011-07-01"), by = "day", length.out = nrow(TBATS_df)),.before = 1)

#export as an excel sheet
write_csv(TBATS_df, 'Submissions/TBATS.csv')
```


#neural network, p=1 P=0, K=2,12
```{r NNETAR, echo=TRUE, message=FALSE, warning=FALSE}
#test diff p and P values, also can add xreg with Fourier term for multiple seasonality

NN_1 <- nnetar(msts_load,
                 p=1,
                 P=0,
                 xreg=fourier(msts_load, K=c(2,12)))

NN_1_forecast <- forecast(NN_1, 
                   h=31,
                   xreg=fourier(msts_load, K=c(2,12),h=31))

#create dataframe /prep for export 
NN_1_df <- data.frame(load = NN_1_forecast$mean)
NN_1_df <- NN_1_df %>% 
  add_column(date = seq(as.Date("2011-07-01"), by = "day", length.out = nrow(NN_1_df)),.before = 1)

#export as an excel sheet
write_csv(NN_1_df, 'Submissions/NN_1.csv')
```

#neural network, p=7 P=0, K=2,12
internet suggested p=7 / a higher p would work better. maybe.
```{r NNETAR, echo=TRUE, message=FALSE, warning=FALSE}
#test diff p and P values, also can add xreg with Fourier term for multiple seasonality

NN_2 <- nnetar(msts_load,
                 p=7,
                 P=0,
                 xreg=fourier(msts_load, K=c(2,12)))

NN_2_forecast <- forecast(NN_2, 
                   h=31,
                   xreg=fourier(msts_load, K=c(2,12),h=31))

#create dataframe /prep for export 
NN_2_df <- data.frame(load = NN_2_forecast$mean)
NN_2_df <- NN_2_df %>% 
  add_column(date = seq(as.Date("2011-07-01"), by = "day", length.out = nrow(NN_2_df)),.before = 1)

#export as an excel sheet
write_csv(NN_2_df, 'Submissions/NN_2.csv')
```

slightly better with higher p value. will attempt slightly different K values

#neural network, p=7 P=0, K=2,6
```{r NNETAR, echo=TRUE, message=FALSE, warning=FALSE}
#test diff p and P values, also can add xreg with Fourier term for multiple seasonality

NN_3 <- nnetar(msts_load,
                 p=7,
                 P=0,
                 xreg=fourier(msts_load, K=c(2,6)))

NN_3_forecast <- forecast(NN_3, 
                   h=31,
                   xreg=fourier(msts_load, K=c(2,6),h=31))

#create dataframe /prep for export 
NN_3_df <- data.frame(load = NN_3_forecast$mean)
NN_3_df <- NN_3_df %>% 
  add_column(date = seq(as.Date("2011-07-01"), by = "day", length.out = nrow(NN_3_df)),.before = 1)

#export as an excel sheet
write_csv(NN_3_df, 'Submissions/NN_3.csv')
```


Should add temp and humidity to improve my best model - NN_2!
#neural network, p=7 P=0, K=2,12 + temp and humidity variables
```{r}
temp_df <- read_excel("temperature.xlsx") %>% 
  mutate(hourly_temp = rowMeans(across(3:26), na.rm = TRUE)) %>% 
  mutate(date = ymd(date)) %>% 
  group_by(date) %>%
  summarise(daily = mean(as.numeric(hourly_temp), na.rm = TRUE)) %>%
  select(daily) %>%
  na.omit(temp)

temp <- temp_df$daily #vector
```

```{r}
humidity_df <- read_excel("relative_humidity.xlsx") %>% 
  mutate(hourly_humidity = rowMeans(across(3:26), na.rm = TRUE)) %>% 
  mutate(date = ymd(date)) %>% 
  group_by(date) %>%
  summarise(daily = mean(as.numeric(hourly_humidity), na.rm = TRUE)) %>%
  select(daily) %>%
  na.omit(humidity)

humidity <- humidity_df$daily #vector
```



```{r NNETAR, echo=TRUE, message=FALSE, warning=FALSE}

NN_4 <- nnetar(msts_load,
                 p=7,
                 P=0,
                 xreg= cbind(fourier(msts_load, K=c(2,12)), temp, humidity))

temp_July <- forecast(auto.arima(temp), h = 31) #forecasting future 31 from the vector
humidity_July <- forecast(auto.arima(humidity), h = 31) #forecasting future 31 from the vector (must have to include below)

NN_4_forecast <- forecast(NN_4, h=31,
                   xreg= cbind(fourier(msts_load, K = c(2,12), h = 31),
                               temp_July$mean,   
                               humidity_July$mean))

#create dataframe /prep for export 
NN_4_df <- data.frame(load = NN_4_forecast$mean)
NN_4_df <- NN_4_df %>% 
  add_column(date = seq(as.Date("2011-07-01"), by = "day", length.out = nrow(NN_4_df)),.before = 1)

#export as an excel sheet
write_csv(NN_4_df, 'Submissions/NN_4.csv')
```

#prophet
```{r}



```


